%!TEX root = ../main.tex
\section{Lab Activity 4}
This lab assignment requires the usage of a \textbf{motor schema} in order to
implement a behaviour-based control for the task of phototaxis, while avoiding
collisions with other objects. The constraints are about maximum wheels
velocities ($15^{-2}m/s$) and that the robot must reach the light source as
fast as possible.

\subsection{Motor Schema Design}
First thing first, we define the potential fields that characterise the environment:
\begin{itemize}
	\item An \emph{attractive} potential field is applied to the light source,
	      with a variation that resembles the concept of ``love'' for type three
	      Breitenberg vehicles. This enables the robot to slow down or even stop
	      when approaching the light source.
	\item A \emph{repulsive} potential field for obstacles is applied, making
	      the robot steer harder away from the obstacle the closer it is.
\end{itemize}

When it comes to motor schemas, they can be summarised through \Cref{tab:ms-lab4}.

\begin{table}[ht]
	\centering
	\begin{tabular}{@{}ll@{}}
		\toprule
		\textbf{Percept Source}   & \textbf{Perceptual Schema}                                                                                                                      \\ \midrule
		Light sensors             & \begin{tabular}[c]{@{}l@{}}Phototaxis performed following  direction (angle) of \\ maximum sensor's value.\end{tabular}                         \\
		Proximity sensors         & \begin{tabular}[c]{@{}l@{}}Collision avoidance performed taking the \\ opposite direction (angle) of the maximum sensor's\\ value.\end{tabular} \\
		Light + Proximity sensors & \begin{tabular}[c]{@{}l@{}}If max light value and max proximity value are below\\ a given threshold, perform a random walk.\end{tabular}        \\ \bottomrule
	\end{tabular}
	\caption{Summary of sources and perception schemas for performing phototaxis with collision avoidance}
	\label{tab:ms-lab4}
\end{table}

For each perceptual schema, we can define the following rules for generating
the resulting vector:
\begin{itemize}
	\item \emph{Phototaxis}: \texttt{length} is equals to the maximum value detected
	      from sensors, while its \texttt{angle} is equals to the angle of the
	      sensor with maximum value.

	\item \emph{Collision Avoidance}: \texttt{length} is equals to the maxim
	      value detected from sensors, while its \texttt{angle} is equals to the
	      opposite angle of the sensor with maximum value.

	\item \emph{Random Walk}: \texttt{length} is equals to 0 when max light value
	      or max proximity value are above thresholds, a constant (0.2) otherwise.
	      The angle is randomly picked each $N$ steps.
\end{itemize}

After generating all vectors, they will be summed up creating a resulting vector
which will be used to apply the corresponding velocities after its conversion
to the differential steering model.

With this approach we have several advantages including a substantial
conciseness when writing the controller, and an overall simplicity when it
comes to behaviours' resulting vectors merging. Challenges arises when having
to produce such vectors. For this kind of task it has been quite
straightforward to assign vector values with directly associating their length
and angle to the sensors. In other scenarios, especially with heterogeneous
sensor kinds and values representation, a Motor Schema architecture can lead to
complexities that could go well beyond the single robot behaviour.

Comparing with the subsumption architecture seen in \Cref{ssec:subsum}, we have
two very distinct and somewhat opposite implementation strategies: with the
subsumption architecture, the design phase is quite easy, but requires careful
consideration in terms of layers interactions and inhibition of lower levels
when taking control during the actual implementation of it. Quite the opposite
happened in a motor schema architecture. The design phase is crucial and
defining potential fields, perceptual schemas and their corresponding vector
mapping is essential in order to have a correct behaviour, while the actual
implementation is trivial. When comes to how well they behave in terms of
\emph{extendibility}, I find the subsumption architecture more flexible. As
said above, a possibile different sensor must be mapped in the common space
defined for all other vectors, making this procedure sometimes challenging
(just think about motor ground sensors that are less in number and yields sort
of boolean values, very different from proximity or light sensors). This
requires, in my opinion, a higher level of complexity when defining the
perceptual schema, with respect to the addition of a new layer in a well
structured subsumption architecture.

A hybrid architecture could be developed picking the best from both
architectures. In my experience, if we keep a subsumption architecture as
simple as possibile, the resulting behaviour is less performing than the motor
schema one. This happens because in the provided implementation of
\Cref{ssec:subsum}, high level behaviours are excluding each other, meaning
that when an obstacle is detected, no matter where the light is, the opposite
direction is taken. This is not true for the MS architecture, while the
contribution of the light (if visible) is still influencing the collision
avoidance behaviour, without the need of additional code thanks to vector
summing. With these consideration in mind, a subsumption architecture could be
designed for all lower level capabilities and for higher level ones that
requires to be exclusively performed (e.g. halting on black spot). For
competing behaviours, we could design a layer (or more) taking advantage of the
MS architecture.
In this way we keep the modularity of the subsumption architecture, while
keeping the effortless behaviour fusing capability provided by the MS
architecture only in some specific layers that requires it.

\subsection{Testing}
For testing, \Cref{ssec:sub-testing} subsumption architecture's test set has
been used. While it is possibile to use the same arena random seed (so same
walls and robot initial positioning), it was not possibile to reset the control
software random seed by means of the simulator. Therefore in this test results
shown in \Cref{tab:ms-results} only a single run is displayed.

\begin{table}[ht]
	\centering
	\begin{tabular}{cccc}
		\textbf{Arena 1} (631098) & \textbf{Arena 2} (859511) & \textbf{Arena 3} (446461) & \textbf{Arena 4} (969235) \\
		\hline
		\textcolor{blue}{4.163}   & \textcolor{blue}{0.284}   & 966                       & 805                       \\
	\end{tabular}
	\caption{Simulation results on 1000 steps runs for the motor schema
		architecture: integer scores indicate then number of steps in order to
		reach the light; floating-point scores (in \textcolor{blue}{blue})
		indicate that the footbot could not reach the light in 1000 steps and
		records the minimum euclidean distance reached from the light. In
		column header, inside parenthesis are contained random seeds (argos) in
		order to produce the arena and robot placement.}
	\label{tab:ms-results}
\end{table}

The motor schema architecture is very sensitive to parameters tuning, and the
tested configuration should not be to consider the optimal one. In fact,
reaching light as fast as possibile is quite challenging to obtain due to the
adjustments performed to the resulting vector. However, the obtained results
can be considered acceptable, especially with respect to the subsumption
architecture. Apart from the first arena, the motor schema architecture yields
very positive results in all other three. When it comes to the actual number of
steps that it takes for the footbot to reach the light, it easily predictable
that it could be a bit higher due to the attractive field of the light: being
the robot forced to stop under the light, the potential field of the light
makes it slow down when in its proximity, meaning that even a hundred steps can
pass for the robot to be precisely under the light and thus accomplishing it
task: this explains the arena two results, if given a bit more steps the robot
would have reached the light (recall the results of arena two for the
subsumption architecture). However the overall behaviour of the motor schema
architecture yielded better results and a good balance between sub-tasks than
the subsumption architecture.
